Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2020-08-03 FDT2020/main-section.tex
chris vs FDT plot updated (leneged out of the plot now)

Consider an instance $I=(A,b)$ of the IP formulation. Define sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume $S(I)\subseteq \{0,1\}^n$ and $P(I)\subseteq [0,1]^n$. For simplicity in the notation we assume an instance $I$ and denote $P(I),S(I),$ and $g(I)$ by $P$, $S$, and $g$ for this section and the next section. Also, for both sections we assume $t=|\spp(x)|$. Without loss of generality we can assume $x_i = 0$ for $i=t+1,\ldots,n$.
\cindy{TODO: Is this something that should be clear to the reader with what they know at this point? Does it need an explanation? Does it need an explanation at the end of this section or the next?}

@@

\begin{figure}[h!]
\begin{subfigure}{.5\textwidth}
\centering
	\includegraphics[width=8cm,scale=1]{"fpvsfdt".png}
	\caption{Dive FDT vs feasbility pump on the instances of PACE 2019 \cite{PACE} with 200 vertices.}
	\label{fpvsfdt}
	\end{subfigure}
	$\quad\;$
	\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=8cm,scale=1]{"fjvsfdt".png}
%	\caption{FDT vs the 2-approximation for TAP\cite{FJ81} on randomly generated extreme points of the cut-LP.}
	\caption{TAP on our random instances: FDT run on LP optimal vs the 2-approximation for TAP\cite{FJ81}.}
	\label{fjvsfdt}
	\end{subfigure}
	\caption{Computational experiments with the FDT algorithm}
	\label{fdtcomp}
\end{figure}
\paragraph{FDT on randomly generated instances of TAP.}
Recall that in the tree-augmentation problem (TAP) we are given a tree $T=(V,E)$, a set of non-tree links $L$ between vertices in $V$ and costs $c\in \mathbb{R}^{L}_{\geq 0}$. A feasible augmentation is $L'\subseteq L$ such that $T+L'$ is 2-edge-connected. In TAP we wish to find the minimum-cost feasible augmentation. The integrality gap of the cut LP for TAP (given in (\ref{eq:cutLP})) is
\begin{equation*}
g(\tap) = \max_{c\in \mathbb{R}^L_{\geq 0}} \frac{\min_{x\in\tap(T,L)} cx}{\min_{x\in\cut(T,L)} cx},
\end{equation*}  
where $TAP(T,L)$ is the feasible set for the IP and $CUT(T,L)$ is the feasible set for the cut LP (relaxation). We know $\frac{3}{2}\leq g(\tap)\leq 2$~\cite{FJ81,32gaptap}. Notice that $\min_{x\in \tap(T,L)}cx$ is a binary IP.

@@

\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c}
			\hline
			& number of edges in $T$ & number of links in $L$ & number of instances $(T,L)$\\ \hline
			 & $6$ & $6$ & $100$ \\ 
			 & $14$ & $28$ & $100$ \\ 
			 & $30$ & $120$ & $100$ \\
			 & $62$ & $496$ & $100$ \\ 
			  & $126$ & $2016$ & $50$ \\  \hline \\
		\end{tabular}\caption{Summary of the randomly generated instances of TAP.} \label{tableTAP}
	\end{small}
\end{table}
\cindy{Please check the edits to the next paragraph carefully to ensure I captured what you were trying to say.}
For all 450 instances in our experiments, running FDT on the LP optimal (fractional extreme point) of the cut LP gave a feasible solution with value at most a factor $\frac{3}{2}$ larger than the LP lower bound.  Such a feasible solution gives an upper bound on the integrality gap $g(\tap)$ of that specific instance of at most $\frac{3}{2}$.
\cindy{There are 450 instances, so you tested one extreme point (the LP optimal) per instance?  I'm assuming that was the LP optimal.}
%when restricted to the instances considered.
In fact, the integrality gap upper bound derived this way was equal to $\frac{3}{2}$ for only one instance.
%the upper bound we provided by our experiment is at $\frac{3}{2}$.
For 425 instances, the integrality gap upper bound was $\frac{4}{3}$, for 22 instances it was $\frac{6}{5}$, for 2 instances it was $\frac{8}{7}$.
\paragraph{Computational comparison between Christofides' algorithm and FDT for 2EC on Carr-Vempala points.}

>>>>

Consider an instance $I=(A,b)$ of the IP formulation. Define sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume $S(I)\subseteq \{0,1\}^n$ and $P(I)\subseteq [0,1]^n$. For simplicity in the notation we assume an instance $I$ and denote $P(I),S(I),$ and $g(I)$ by $P$, $S$, and $g$ for this section and the next section. Also, for both sections we assume $t=|\spp(x)|$. Without loss of generality we can assume $x_i = 0$ for $i=t+1,\ldots,n$.
\cindy{TODO: Is this something that should be clear to the reader with what they know at this point? Does it need an explanation? Does it need an explanation at the end of this section or the next?}\arash{this is just a relabeling of the indices.}

@@

\begin{figure}[h!]
\begin{subfigure}{.5\textwidth}
\centering
	\includegraphics[width=8cm,scale=1]{"fpvsfdt".png}
	\caption{Dive FDT vs feasbility pump on the instances of PACE 2019 \cite{PACE} with 200 vertices.}
	\label{fpvsfdt}
	\end{subfigure}
	$\quad\;$
	\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=8cm,scale=1]{"fjvsfdt".png}
%	\caption{FDT vs the 2-approximation for TAP\cite{FJ81} on randomly generated extreme points of the cut-LP.}
	\caption{TAP on our random instances: FDT run on LP optimal vs the 2-approximation for TAP\cite{FJ81}.}
	\label{fjvsfdt}
	\end{subfigure}
	\caption{Computational experiments with the FDT algorithm}
	\label{fdtcomp}
\end{figure}
\paragraph{FDT on randomly generated instances of TAP.}
Recall that in the tree-augmentation problem (TAP) we are given a tree $T=(V,E)$, a set of non-tree links $L$ between vertices in $V$ and costs $c\in \mathbb{R}^{L}_{\geq 0}$. A feasible augmentation is $L'\subseteq L$ such that $T+L'$ is 2-edge-connected. In TAP we wish to find the minimum-cost feasible augmentation. The integrality gap of the cut LP for TAP (given in (\ref{eq:cutLP})) is
\begin{equation*}
g(\tap) = \max_{c\in \mathbb{R}^L_{\geq 0}} \frac{\min_{x\in\tap(T,L)} cx}{\min_{x\in\cut(T,L)} cx},
\end{equation*}  
where $TAP(T,L)$ is the feasible set for the IP and $CUT(T,L)$ is the feasible set for the cut LP (relaxation). \arash{Did you intentionally change the TAP and CUT module? They print as italic now as opposed to how they look in $\tap(T,L)$ and $\cut(T,L)$} We know $\frac{3}{2}\leq g(\tap)\leq 2$~\cite{FJ81,32gaptap}. Notice that $\min_{x\in \tap(T,L)}cx$ is a binary IP.

@@

\begin{table}[h!]
	\begin{small}
		\centering
		\begin{tabular}{c c c c}
			\hline
			& number of edges in $T$ & number of links in $L$ & number of instances $(T,L)$\\ \hline
			 & $6$ & $6$ & $100$ \\ 
			 & $14$ & $28$ & $100$ \\ 
			 & $30$ & $120$ & $100$ \\
			 & $62$ & $496$ & $100$ \\ 
			  & $126$ & $2016$ & $50$ \\  \hline \\
		\end{tabular}\caption{Summary of the randomly generated instances of TAP.} \label{tableTAP}
	\end{small}
\end{table}
\cindy{Please check the edits to the next paragraph carefully to ensure I captured what you were trying to say.}
For all 450 instances in our experiments, running FDT on the LP optimal (fractional extreme point) of the cut LP gave a feasible solution with value at most a factor $\frac{3}{2}$ larger than the LP lower bound.  Such a feasible solution gives an upper bound on the integrality gap $g(\tap)$ of that specific instance of at most $\frac{3}{2}$.
\cindy{There are 450 instances, so you tested one extreme point (the LP optimal) per instance?  I'm assuming that was the LP optimal.}
%when restricted to the instances considered.
In fact, the integrality gap upper bound derived this way was equal to $\frac{3}{2}$ for only one instance.
%the upper bound we provided by our experiment is at $\frac{3}{2}$.
For 425 instances, the integrality gap upper bound was $\frac{4}{3}$, for 22 instances it was $\frac{6}{5}$, for 2 instances it was $\frac{8}{7}$.
\arash{it's all correct, I'm running the experiments again to get the running time so these numbers might change.}
\paragraph{Computational comparison between Christofides' algorithm and FDT for 2EC on Carr-Vempala points.}
